---
title: "Lab 02 - Plastic waste"
author: "Elayna Seago"
date: "1/25/22"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

The continents show quite a lot of similarity, for all continents people in the majority of countries produce less than 1kg of plastic waste per year. The exception is Trinidad and Tobago, shown on the graph of North America, where people produce more than 3.5kg of plastic waste per year.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap >3.5)

ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~continent)

ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_density()

ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap,
                     color = continent,
                     fill = continent)) +
  geom_density(alpha = .7)

```

### Exercise 2

Defining specifications in aesthetics mapping changes how variables are mapped onto the graph while defining specifications in geom_density changes elements of the graph (regardless of what variable they are associated with).
```{r plastic-waste-density}
ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap,
                     color = continent,
                     fill = continent)) +
  geom_density(alpha = .3)

ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap))+
  geom_boxplot()
```

### Exercise 3

The box plots show each outlier more distinctly with individual dots per outlier. The violin plot makes it impossible to miss an outlier.The violin plot also shows the density of data at each value of plastic waste per capita.

```{r plastic-waste-violin}
ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap))+
  geom_violin()
```

### Exercise 4

There is a positive relationship between plastic waste per capita and mismanaged plastic waste per capita. This makes sense because the more waste you have the more waste you can mismanage.

```{r scatter plot of plastic-waste-per-cap and mismanaged-plastic-waste}
ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

### Exercise 4.2
There are some distinctions. Europe seems to have a lower rate of mismanaged plastic per capita given the amount of plastic waste they have per capita. North America is also doing well barring Trinidad and Tobago. 

```{r}
ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point()
```

### Exercise 4.3
The relationship between these pairs looks fairly similar to me.

```{r}
ggplot(data = plastic_waste,
       mapping = aes(x = total_pop, y = plastic_waste_per_cap, color = continent))+
  geom_point()


ggplot(data = plastic_waste,
       mapping = aes(x = coastal_pop, y = plastic_waste_per_cap, color = continent))+
  geom_point()

```



### Exercise 5

This graph plots the relationship between the proportion of a country's population that is coastal and the amount of plastic waste per capita in that country. There is also a smoothing line with standard error bars that shows the trend. There appears to be a positive relationship between proportion of coastal population and plastic waste, though the relationship doesn't really look that linear.

```{r plastic-waste-mismanaged}

ggplot(data = plastic_waste%>%
         filter(plastic_waste_per_cap<3),
       mapping = aes(x =(coastal_pop/total_pop), y = plastic_waste_per_cap, color = continent))+
  geom_point()+
  geom_smooth(inherit.aes = FALSE, aes(x =(coastal_pop/total_pop), y = plastic_waste_per_cap))+
  labs(title = "Plastic waste vs Coastal population proportion",
       subtitle = "by continent",x = "Coastal population proportion (coastal/total population)")

```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



