---
title: "Lab 02 - Plastic waste"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

The continents show quite a lot of similarity, for all continents people in the majority of countries produce less than 1kg of plastic waste per year. The exception is Trinidad and Tobago, shown on the graph of North America, where people produce more than 3.5kg of plastic waste per year.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap >3.5)

ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~continent)

ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_density()

ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap,
                     color = continent,
                     fill = continent)) +
  geom_density(alpha = .7)

```

### Exercise 2

Defining specifications in aesthetics mapping changes how variables are mapped onto the graph while defining specifications in geom_density changes elements of the graph (regardless of what variable they are associated with).
```{r plastic-waste-density}
ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap,
                     color = continent,
                     fill = continent)) +
  geom_density(alpha = .3)

ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap))+
  geom_boxplot()
```

### Exercise 3

The box plots show each outlier more distinctly with individual dots per outlier. The violin plot makes it impossible to miss an outlier.The violin plot also shows the density of data at each value of plastic waste per capita.

```{r}
ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap))+
  geom_violin()
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



